---
title: "Automated_immunology_analysis_Rules"
author: "Carlos Eduardo Madureira Trufen"
date: "6/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Load libraries
```{r, message=FALSE, warning=FALSE, cache=TRUE}
library(bestNormalize, warn.conflicts = FALSE)
library(caret, warn.conflicts = FALSE)
library(corrplot, warn.conflicts = FALSE)
library(clusterCrit, warn.conflicts = FALSE)
library(CytoML, warn.conflicts = FALSE)
library(DiffSOM, warn.conflicts = FALSE)
library(doParallel, warn.conflicts = FALSE)
library(EmbedSOM, warn.conflicts = FALSE)
library(FactoMineR, warn.conflicts = FALSE)
library(factoextra, warn.conflicts = FALSE)
library(flowClean, warn.conflicts = FALSE)
library(flowCore, warn.conflicts = FALSE)
library(flowClust, warn.conflicts = FALSE)
library(flowStats, warn.conflicts = FALSE)
library(flowViz, warn.conflicts = FALSE)
library(flowWorkspace, warn.conflicts = FALSE)
library(FlowSOM, warn.conflicts = FALSE)
library(furrr, warn.conflicts = FALSE)
library(ggcyto, warn.conflicts = FALSE)
library(HDclassif, warn.conflicts = FALSE)
library(kohonen, warn.conflicts = FALSE)
library(microbenchmark, warn.conflicts = FALSE)
library(parallel, warn.conflicts = FALSE)
library(parallelMap, warn.conflicts = FALSE)
library(profvis, warn.conflicts = FALSE)
library(RColorBrewer, warn.conflicts = FALSE)
library(Rcpp11, warn.conflicts = FALSE)
library(tictoc, warn.conflicts = FALSE)
library(tidyverse, warn.conflicts = FALSE)
library(yasomi, warn.conflicts = FALSE)
# source("~/PRIMUS/home/trufenc/new_SOM_functions.R")
source("/run/user/1002/gvfs/smb-share:server=primus.img.cas.cz,share=home/trufenc/new_SOM_functions.R")
```

```{r}
scale_to_range_0_1 <- function(x, ...){(x - min(x, ...)) / (max(x, ...) - min(x, ...))}
```

path to files
```{r path_to_files, cache=TRUE}
# path_to_files <- "~/PRIMUS/data/83_BIOINFORMATICS/Carlos"
path_to_files <- "/run/user/1002/gvfs/smb-share:server=primus.img.cas.cz,share=data/83_BIOINFORMATICS/Carlos"
# path_to_files <- "smb://primus.img.cas.cz/data/83_BIOINFORMATICS/Carlos"
```

```{r}
path_to_control_files_1 <- "/run/user/1002/gvfs/smb-share:server=primus.img.cas.cz,share=data/83_IMPC_DATA/Immunology/CCP_Phenotyping_Immunology_Data_JB/CCP_cohorts_2018/CCP_cohorts_2018_12/2018_12_27_C067/2018_12_27_PA/C067_cA1556_gates"

path_to_control_files_2 <- "/run/user/1002/gvfs/smb-share:server=primus.img.cas.cz,share=data/83_IMPC_DATA/Immunology/CCP_Phenotyping_Immunology_Data_JB/CCP_cohorts_2019/CCP_cohorts_2019_10/2019_10_29_C108/2019_10_29_PA/gates_c90692"

path_to_control_files_3 <- "/run/user/1002/gvfs/smb-share:server=primus.img.cas.cz,share=data/83_IMPC_DATA/Immunology/CCP_Phenotyping_Immunology_Data_JB/CCP_cohorts_2019/CCP_cohorts_2019_10/2019_10_29_C108/2019_10_29_PA/gates_c90695"

path_to_ko_files_1 <- "/run/user/1002/gvfs/smb-share:server=primus.img.cas.cz,share=data/83_IMPC_DATA/Immunology/CCP_Phenotyping_Immunology_Data_JB/CCP_cohorts_2019/CCP_cohorts_2019_10/2019_10_29_C108/2019_10_29_PA/gates_tgE8038"

path_to_ko_files_2 <- "/run/user/1002/gvfs/smb-share:server=primus.img.cas.cz,share=data/83_IMPC_DATA/Immunology/CCP_Phenotyping_Immunology_Data_JB/CCP_cohorts_2019/CCP_cohorts_2019_10/2019_10_29_C108/2019_10_29_PA/gates_tgE8040"
```

```{r}
path_to_files <- list(path_to_control_files_1,
                      path_to_control_files_2,
                      path_to_control_files_3)

path_to_ko_files <- list(path_to_ko_files_1,
                           path_to_ko_files_2)
```

```{r files, cache=TRUE}
files <- path_to_files %>% purrr::map(list.files, pattern = ".fcs", full.names = TRUE)
ko_files <- path_to_ko_files %>% purrr::map(list.files, pattern = ".fcs", full.names = TRUE)
```

```{r file_names, cache=TRUE}
file_names <- path_to_files %>% purrr::map(list.files, pattern = ".fcs")
ko_files_names <- path_to_ko_files %>% purrr::map(list.files, pattern = ".fcs")
```

```{r}
panel_A_classes <- c(
  "_clog\\+.fcs",
  "_clog\\-_cells\\-.fcs",
  "_S1\\-.fcs",
  "_dead.fcs",
  "_S2\\-.fcs",
  "gdTCR_Eff_Klrg1\\-.fcs",
  "gdTCR_Eff_Klrg1\\+.fcs",
  "gdTCR_Res_Klrg1\\-.fcs",
  "gdTCR_Res_Klrg1\\+.fcs",
  "gdTCR_neg_Klrg1\\-.fcs",
  "gdTCR_neg_Klrg1\\+.fcs",
  "NK_Eff_Klrg1\\-.fcs",
  "NK_Eff_Klrg1\\+.fcs",
  "NK_Res_Klrg1\\-.fcs",
  "NK_Res_Klrg1\\+.fcs",
  "NK_neg_Klrg1\\+\\-.fcs",
  "NK_neg_Klrg1\\+.fcs",
  #"_CD5\\+\\-.fcs",
  #"_CD4\\+ NKT\\+ or CD4\\- NKT\\+\\-.fcs",
  "CD4\\-NKT_Eff_Klrg1\\-.fcs",
  "CD4\\-NKT_Eff_Klrg1\\+.fcs",
  "CD4\\-NKT_Res_Klrg1\\-.fcs",
  "CD4\\-NKT_Res_Klrg1\\+.fcs",
  "CD4\\-NKT_neg_Klrg1\\-.fcs",
  "CD4\\-NKT_neg_Klrg1\\+.fcs",
  "CD4\\+NKT_Eff_Klrg1\\-.fcs",
  "CD4\\+NKT_Eff_Klrg1\\+.fcs",
  "CD4\\+NKT_Res_Klrg1\\-.fcs",
  "CD4\\+NKT_Res_Klrg1\\+.fcs",
  "CD4\\+NKT_neg_Klrg1\\-.fcs",
  "CD4\\+NKT_neg_Klrg1\\+.fcs",
  "DN .fcs",
  "Tregs_Eff_Klrg1\\-.fcs",
  "Tregs_Eff_Klrg1\\+.fcs",
  "Tregs_Res_Klrg1\\-.fcs",
  "Tregs_Res_Klrg1\\+.fcs",
  "Tregs_neg_Klrg1\\-.fcs",
  "Tregs_neg_Klrg1\\+.fcs",
  "Th_Eff_Klrg1\\-.fcs",
  "Th_Eff_Klrg1\\+.fcs",
  "Th_Res_Klrg1\\-.fcs",
  "Th_Res_Klrg1\\+.fcs",
  "Th_neg_Klrg1\\-.fcs",
  "Th_neg_Klrg1\\+.fcs",
  "CD8_Naive_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8_Naive_Klrg1\\-_VM Tcells.fcs",
  "CD8_Naive_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8_Naive_Klrg1\\+_VM Tcells.fcs",
  "CD8_Eff_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8_Eff_Klrg1\\-_VM Tcells.fcs",
  "CD8_Eff_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8_Eff_Klrg1\\+_VM Tcells.fcs",
  "CD8_Res_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8_Res_Klrg1\\-_VM Tcells.fcs",
  "CD8_Res_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8_Res_Klrg1\\+_VM Tcells.fcs",
  "CD8_neg_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8_neg_Klrg1\\-_VM Tcells.fcs",
  "CD8_neg_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8_neg_Klrg1\\+_VM Tcells.fcs"
)
```

```{r}
indices_1 <- purrr::map(seq_along(panel_A_classes), function(i) files[[1]] %>% str_which(pattern = panel_A_classes[i])) %>% unlist()
files_to_load_1 <- files[[1]][indices_1]
files_to_load_names_1 <- file_names[[1]][indices_1]
```

# Second class assignment, more refined - gates_c90692
```{r}
panel_A_classes_2 <- c(
  "clog\\-\\-.fcs",
  "cells\\-.fcs",
  "S1\\-.fcs",
  "dead.fcs",
  "S2\\-.fcs",
  "gd_Eff_Klrg1\\+\\-.fcs",
  "gd_Eff_Klrg1\\+.fcs",
  "gd_Res_Klrg1\\+\\-.fcs",
  "gd_Res_Klrg1\\+.fcs",
  "gd_neg_Klrg1\\+\\-.fcs",
  "gd_neg_Klrg1\\+.fcs",
  "NK_Eff_Klrg1\\+\\-.fcs",
  "NK_Eff_Klrg1\\+.fcs",
  "NK_Res_Klrg1\\+\\-.fcs",
  "NK_Res_Klrg1\\+.fcs",
  "NK_neg_Klrg1\\+\\-.fcs",
  "NK_neg_Klrg1\\+.fcs",
  "CD4\\-NKT_Eff_Klrg1\\+\\-.fcs",
  "CD4\\-NKT_Eff_Klrg1\\+.fcs",
  "CD4\\-NKT_Res_Klrg1\\+\\-.fcs",
  "CD4\\-NKT_Res_Klrg1\\+.fcs",
  "CD4\\-NKT_neg_Klrg1\\+\\-.fcs",
  "CD4\\-NKT_neg_Klrg1\\+.fcs",
  "CD4\\+NKT_Eff_Klrg1\\+\\-.fcs",
  "CD4\\+NKT_Eff_Klrg1\\+.fcs",
  "CD4\\+NKT_Res_Klrg1\\+\\-.fcs",
  "CD4\\+NKT_Res_Klrg1\\+.fcs",
  "CD4\\+NKT_neg_Klrg1\\+\\-.fcs",
  "CD4\\+NKT_neg_Klrg1\\+.fcs",
  "DN .fcs",
  "Treg_Eff_Klrg1\\+\\-.fcs",
  "Treg_Eff_Klrg1\\+.fcs",
  "Treg_Res_Klrg1\\+\\-.fcs",
  "Treg_Res_Klrg1\\+.fcs",
  "Treg_neg_Klrg1\\+\\-.fcs",
  "Treg_neg_Klrg1\\+.fcs",
  "Th_Eff_Klrg1\\+\\-.fcs",
  "Th_Eff_Klrg1\\+.fcs",
  "Th_Res_Klrg1\\+\\-.fcs",
  "Th_Res_Klrg1\\+.fcs",
  "Th_neg_Klrg1\\+\\-.fcs",
  "Th_neg_Klrg1\\+.fcs",
  "CD8\\+_Naive_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8\\+_Naive_Klrg1\\-_VM Tcells.fcs",
  "CD8\\+_Naive_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8\\+_Naive_Klrg1\\+_VM Tcells.fcs",
  "CD8\\+_Eff_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8\\+_Eff_Klrg1\\-_VM Tcells.fcs",
  "CD8\\+_Eff_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8\\+_Eff_Klrg1\\+_VM Tcells.fcs",
  "CD8\\+_Res_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8\\+_Res_Klrg1\\-_VM Tcells.fcs",
  "CD8\\+_Res_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8\\+_Res_Klrg1\\+_VM Tcells.fcs",
  "CD8\\+_neg_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8\\+_neg_Klrg1\\-_VM Tcells.fcs",
  "CD8\\+_neg_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8\\+_neg_Klrg1\\+_VM Tcells.fcs"
)
```

```{r}
indices_2 <- purrr::map(seq_along(panel_A_classes_2), function(i) files[[2]] %>% str_which(pattern = panel_A_classes_2[i])) %>% unlist()
files_to_load_2 <- files[[2]][indices_2]
files_to_load_names_2 <- file_names[[2]][indices_2]

files_to_load_2[files_to_load_2 %>% str_which(pattern = "CD8\\+_VM Tcells\\-")] <- NA
files_to_load_2 <- files_to_load_2[!is.na(files_to_load_2)]
```

# Second class assignment, more refined - gates_c90695
```{r}
panel_A_classes_3 <- c(
  "clog\\-\\-.fcs",
  "clog\\-_cells\\-.fcs",
  "S1\\-.fcs",
  "dead.fcs",
  "S2\\-.fcs",
  "gd_Eff_Klrg1\\+\\-.fcs",
  "gd_Eff_Klrg1\\+.fcs",
  "gd_Res_Klrg1\\+\\-.fcs",
  "gd_Res_Klrg1\\+.fcs",
  "gd_neg_Klrg1\\+\\-.fcs",
  "gd_neg_Klrg1\\+.fcs",
  "NK_Eff_Klrg1\\+\\-.fcs",
  "NK_Eff_Klrg1\\+.fcs",
  "NK_Res_Klrg1\\+\\-.fcs",
  "NK_Res_Klrg1\\+.fcs",
  "NK_neg_Klrg1\\+\\-.fcs",
  "NK_neg_Klrg1\\+.fcs",
  "CD4\\-NKT_Eff_Klrg1\\+\\-.fcs",
  "CD4\\-NKT_Eff_Klrg1\\+.fcs",
  "CD4\\-NKT_Res_Klrg1\\+\\-.fcs",
  "CD4\\-NKT_Res_Klrg1\\+.fcs",
  "CD4\\-NKT_neg_Klrg1\\+\\-.fcs",
  "CD4\\-NKT_neg_Klrg1\\+.fcs",
  "CD4\\+NKT_Eff_Klrg1\\+\\-.fcs",
  "CD4\\+NKT_Eff_Klrg1\\+.fcs",
  "CD4\\+NKT_Res_Klrg1\\+\\-.fcs",
  "CD4\\+NKT_Res_Klrg1\\+.fcs",
  "CD4\\+NKT_neg_Klrg1\\+\\-.fcs",
  "CD4\\+NKT_neg_Klrg1\\+.fcs",
  "DN .fcs",
  "Treg_Eff_Klrg1\\+\\-.fcs",
  "Treg_Eff_Klrg1\\+.fcs",
  "Treg_Res_Klrg1\\+\\-.fcs",
  "Treg_Res_Klrg1\\+.fcs",
  "Treg_neg_Klrg1\\+\\-.fcs",
  "Treg_neg_Klrg1\\+.fcs",
  "Th_Eff_Klrg1\\+\\-.fcs",
  "Th_Eff_Klrg1\\+.fcs",
  "Th_Res_Klrg1\\+\\-.fcs",
  "Th_Res_Klrg1\\+.fcs",
  "Th_neg_Klrg1\\+\\-.fcs",
  "Th_neg_Klrg1\\+.fcs",
  "CD8\\+_Naive_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8\\+_Naive_Klrg1\\-_VM Tcells.fcs",
  "CD8\\+_Naive_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8\\+_Naive_Klrg1\\+_VM Tcells.fcs",
  "CD8\\+_Eff_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8\\+_Eff_Klrg1\\-_VM Tcells.fcs",
  "CD8\\+_Eff_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8\\+_Eff_Klrg1\\+_VM Tcells.fcs",
  "CD8\\+_Res_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8\\+_Res_Klrg1\\-_VM Tcells.fcs",
  "CD8\\+_Res_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8\\+_Res_Klrg1\\+_VM Tcells.fcs",
  "CD8\\+_neg_Klrg1\\-_VM Tcells\\-.fcs",
  "CD8\\+_neg_Klrg1\\-_VM Tcells.fcs",
  "CD8\\+_neg_Klrg1\\+_VM Tcells\\-.fcs",
  "CD8\\+_neg_Klrg1\\+_VM Tcells.fcs"
)
```

```{r}
indices_3 <- purrr::map(seq_along(panel_A_classes_3), function(i) files[[3]] %>% str_which(pattern = panel_A_classes_3[i])) %>% unlist()
files_to_load_3 <- files[[3]][indices_3]
files_to_load_names_3 <- file_names[[3]][indices_3]
```

All the classes for panel A
```{r}
panel_A_classes_v <- c(
  "clog+",
  "cells-",
  "S1-",
  "dead",
  "S2-",
  "gdTCR_Eff_Klrg1-",
  "gdTCR_Eff_Klrg1+",
  "gdTCR_Res_Klrg1-",
  "gdTCR_Res_Klrg1+",
  "gdTCR_neg_Klrg1-",
  "gdTCR_neg_Klrg1+",
  "NK_Eff_Klrg1-",
  "NK_Eff_Klrg1+",
  "NK_Res_Klrg1-",
  "NK_Res_Klrg1+",
  "NK_neg_Klrg1+-",
  "NK_neg_Klrg1+",
  "CD4-NKT_Eff_Klrg1-",
  "CD4-NKT_Eff_Klrg1+",
  "CD4-NKT_Res_Klrg1-",
  "CD4-NKT_Res_Klrg1+",
  "CD4-NKT_neg_Klrg1-",
  "CD4-NKT_neg_Klrg1+",
  "CD4+NKT_Eff_Klrg1-",
  "CD4+NKT_Eff_Klrg1+",
  "CD4+NKT_Res_Klrg1-",
  "CD4+NKT_Res_Klrg1+",
  "CD4+NKT_neg_Klrg1-",
  "CD4+NKT_neg_Klrg1+",
  "DN",
  "Tregs_Eff_Klrg1-",
  "Tregs_Eff_Klrg1+",
  "Tregs_Res_Klrg1-",
  "Tregs_Res_Klrg1+",
  "Tregs_neg_Klrg1-",
  "Tregs_neg_Klrg1+",
  "Th_Eff_Klrg1-",
  "Th_Eff_Klrg1+",
  "Th_Res_Klrg1-",
  "Th_Res_Klrg1+",
  "Th_neg_Klrg1-",
  "Th_neg_Klrg1+",
  "CD8_Naive_Klrg1-_VM Tcells-",
  "CD8_Naive_Klrg1-_VM Tcells",
  "CD8_Naive_Klrg1+_VM Tcells-",
  "CD8_Naive_Klrg1+_VM Tcells",
  "CD8_Eff_Klrg1-_VM Tcells-",
  "CD8_Eff_Klrg1-_VM Tcells",
  "CD8_Eff_Klrg1+_VM Tcells-",
  "CD8_Eff_Klrg1+_VM Tcells",
  "CD8_Res_Klrg1-_VM Tcells-",
  "CD8_Res_Klrg1-_VM Tcells",
  "CD8_Res_Klrg1+_VM Tcells-",
  "CD8_Res_Klrg1+_VM Tcells",
  "CD8_neg_Klrg1-_VM Tcells-",
  "CD8_neg_Klrg1-_VM Tcells",
  "CD8_neg_Klrg1+_VM Tcells-",
  "CD8_neg_Klrg1+_VM Tcells"
)
```

```{r}
ko_indices_1 <- purrr::map(seq_along(panel_A_classes_3), function(i) ko_files[[1]] %>% str_which(pattern = panel_A_classes_3[i])) %>% unlist()
ko_files_to_load_1 <- ko_files[[1]][ko_indices_1]
ko_files_to_load_names_1 <- ko_files_names[[1]][ko_indices_1]
```

```{r}
ko_indices_2 <- purrr::map(seq_along(panel_A_classes_3), function(i) ko_files[[2]] %>% str_which(pattern = panel_A_classes_3[i])) %>% unlist()
ko_files_to_load_2 <- ko_files[[2]][ko_indices_2]
ko_files_to_load_names_2 <- ko_files_names[[2]][ko_indices_2]
```

# Load files and pre-procees them ()
```{r pre_process_data}
files_to_load <- c(files_to_load_1, files_to_load_2, ko_files_to_load_1, ko_files_to_load_2)
files_to_load[files_to_load %>% str_which(pattern = "CD8\\+_VM Tcells\\-")] <- NA
files_to_load <- files_to_load[!is.na(files_to_load)]

test_files <- c(files_to_load_3)

diff_som_1 <- files_to_load %>% load_fcs_files(n_cells = 1e20, filter = TRUE, toFilter = c(1:6), scale = FALSE, transform = TRUE, toTransform = c(1:18))
test_diff_som <- test_files %>% load_fcs_files(n_cells = 1e20, filter = TRUE, toFilter = c(1:6), scale = FALSE, transform = TRUE, toTransform = c(1:18))

pre_process_data <- function(diff_som, to_transform = c(1:18)){
  # diff_som$fs$data <- diff_som$fs$data %>% purrr::set_names(str_remove(string = names(diff_som$fs$data), pattern = "FJComp."))
  diff_som_df <- diff_som$fs$data %>% as.data.frame() %>% dplyr::select(-c(Time, File, File_scattered))
  normally_dstributed <- diff_som_df %>% purrr::map(bestNormalize::orderNorm)
  normally_dstributed_df <- purrr::map(seq_along(normally_dstributed), function(i) normally_dstributed[[i]]$x.t) %>% purrr::set_names(colnames(diff_som_df)) %>% purrr::map(scale, center = TRUE, scale = TRUE) %>% purrr::map(scale_to_range_0_1) %>% bind_cols()
  diff_som$fs$data[, to_transform] <- normally_dstributed_df %>% as.matrix()
  return(diff_som)
}

diff_som_1 <- diff_som_1 %>% pre_process_data()
test_diff_som <- test_diff_som %>% pre_process_data()
```

# fcs data df
```{r build_fcs_data_df}
# cell_types <- files_to_load_names %>% str_remove(pattern = ".fcs") %>% str_remove(pattern = "export_Specimen_..._c....._..._..._")
cell_types_control <- c(files_to_load_names_1, files_to_load_names_2) %>% 
  str_remove(pattern = ".fcs") %>% 
  str_remove(pattern = "export_Specimen_..._c....._..._..._")

cell_types_ko <- c(ko_files_to_load_names_1, ko_files_to_load_names_2) %>% 
  str_remove(pattern = ".fcs") %>% 
  str_remove(pattern = "export_Specimen_..._tg....._..._..._")

cell_types <- c(cell_types_control, cell_types_ko)
cell_types[cell_types %>% str_which(pattern = "CD8\\+_VM Tcells\\-")] <- NA
cell_types <- cell_types[!is.na(cell_types)]

build_fcs_data_df <- function(cell_types, diff_som) {
  
  cell_types_df <- cell_types %>%
    as.data.frame() %>%
    rownames_to_column() %>%
    dplyr::rename("cell_types" = ".", "File" = "rowname") %>%
    mutate(File = File %>% as.numeric())

  cell_types_df <- cell_types_df %>%
    mutate(cell_types = cell_types %>% recode(
      "clog--" = "clog+",
      "clog-_cells-" = "cells-",
      "cells_S1-" = "S1-",
      "S2_dead" = "dead",
      "S1_S2-" = "S2-",
      "NK_Eff_Klrg1+-" = "NK_Eff_Klrg1-",
      "NK_Res_Klrg1+-" = "NK_Res_Klrg1-",
      "NK_neg_Klrg1+-" = "NK_neg_Klrg1-",
      "gd_Eff_Klrg1+" = "gdTCR_Eff_Klrg1+",
      "gd_Eff_Klrg1+-" = "gdTCR_Eff_Klrg1-",
      "gd_Res_Klrg1+" = "gdTCR_Res_Klrg1+",
      "gd_Res_Klrg1+-" = "gdTCR_Res_Klrg1-",
      "gd_neg_Klrg1+" = "gdTCR_neg_Klrg1+",
      "gd_neg_Klrg1+-" = "gdTCR_neg_Klrg1-",
      "CD4-NKT_Eff_Klrg1+-" = "CD4-NKT_Eff_Klrg1-",
      "CD4-NKT_Res_Klrg1+-" = "CD4-NKT_Res_Klrg1-",
      "CD4-NKT_neg_Klrg1+-" = "CD4-NKT_neg_Klrg1-",
      "CD4+NKT_Eff_Klrg1+-" = "CD4+NKT_Eff_Klrg1-",
      "CD4+NKT_Res_Klrg1+-" = "CD4+NKT_Res_Klrg1-",
      "CD4+NKT_neg_Klrg1+-" = "CD4+NKT_neg_Klrg1-",
      "Th_Eff_Klrg1+-" = "Th_Eff_Klrg1-",
      "Th_Res_Klrg1+-" = "Th_Res_Klrg1-",
      "Th_neg_Klrg1+-" = "Th_neg_Klrg1-",
      "Treg_Eff_Klrg1+-" = "Tregs_Eff_Klrg1-",
      "Treg_Res_Klrg1+-" = "Tregs_Res_Klrg1-",
      "Treg_neg_Klrg1+-" = "Tregs_neg_Klrg1-",
      "Treg_Eff_Klrg1+" = "Tregs_Eff_Klrg1+",
      "Treg_Res_Klrg1+" = "Tregs_Res_Klrg1+",
      "Treg_neg_Klrg1+" = "Tregs_neg_Klrg1+",
      "CD8_Naive_Klrg1-_VM Tcells" = "CD8_Naive_Klrg1-_VM Tcells+",
      "CD8_Naive_Klrg1+_VM Tcells" = "CD8_Naive_Klrg1+_VM Tcells+",
      "CD8_Eff_Klrg1-_VM Tcells" = "CD8_Eff_Klrg1-_VM Tcells+",
      "CD8_Eff_Klrg1+_VM Tcells" = "CD8_Eff_Klrg1+_VM Tcells+",
      "CD8_Res_Klrg1-_VM Tcells" = "CD8_Res_Klrg1-_VM Tcells+",
      "CD8_Res_Klrg1+_VM Tcells" = "CD8_Res_Klrg1+_VM Tcells+",
      "CD8_neg_Klrg1-_VM Tcells" = "CD8_neg_Klrg1-_VM Tcells+",
      "CD8_neg_Klrg1+_VM Tcells" = "CD8_neg_Klrg1+_VM Tcells+",
      "CD8+_Eff_Klrg1-_VM Tcells-" = "CD8_Eff_Klrg1-_VM Tcells-",
      "CD8+_Eff_Klrg1+_VM Tcells" = "CD8_Eff_Klrg1+_VM Tcells+",
      "CD8+_Eff_Klrg1+_VM Tcells-" = "CD8_Eff_Klrg1+_VM Tcells-",
      "CD8+_Naive_Klrg1-_VM Tcells" = "CD8_Naive_Klrg1-_VM Tcells+",
      "CD8+_Naive_Klrg1-_VM Tcells-" = "CD8_Naive_Klrg1-_VM Tcells-",
      "CD8+_Naive_Klrg1+_VM Tcells" = "CD8_Naive_Klrg1+_VM Tcells+",
      "CD8+_Naive_Klrg1+_VM Tcells-" = "CD8_Naive_Klrg1+_VM Tcells-",
      "CD8+_neg_Klrg1-_VM Tcells" = "CD8_neg_Klrg1-_VM Tcells+",
      "CD8+_neg_Klrg1-_VM Tcells-" = "CD8_neg_Klrg1-_VM Tcells-",
      "CD8+_neg_Klrg1+_VM Tcells" = "CD8_neg_Klrg1+_VM Tcells+",
      "CD8+_neg_Klrg1+_VM Tcells-" = "CD8_neg_Klrg1+_VM Tcells-",
      "CD8+_Res_Klrg1-_VM Tcells" = "CD8_Res_Klrg1-_VM Tcells+",
      "CD8+_Res_Klrg1-_VM Tcells-" = "CD8_Res_Klrg1-_VM Tcells-",
      "CD8+_Res_Klrg1+_VM Tcells" = "CD8_Res_Klrg1+_VM Tcells+",
      "CD8+_Res_Klrg1+_VM Tcells-" = "CD8_Res_Klrg1+_VM Tcells-"
    ))

  manual_gating <- diff_som$fs$data %>%
    as.data.frame() %>%
    full_join(cell_types_df) %>%
    na.omit() %>%
    dplyr::select(cell_types) %>%
    unlist() %>%
    as.character() %>%
    factor()

  facs_data_df <- diff_som$fs$data %>%
    as.data.frame() %>%
    dplyr::select(c(1:19))
  # facs_data_df <- diff_som$fs$data %>% as.data.frame() %>% dplyr::select(to_transform)

  facs_data_df <- facs_data_df %>%
    purrr::set_names(str_remove(string = names(facs_data_df), pattern = "FJComp."))

  facs_data_df <- facs_data_df %>%
    mutate(cell_type = manual_gating)
  
  return(facs_data_df)
}

facs_data_df <- diff_som_1 %>% build_fcs_data_df(cell_types = cell_types)

facs_data_df <- facs_data_df %>% 
  mutate(first_gate = case_when(!(cell_type %in% c("clog+", "clog-_cells-", "S1-", "S2-", "dead")) ~ "live"))
  
facs_data_df <- facs_data_df %>% 
  mutate(first_gate = if_else(is.na(first_gate), true = "not_live", false = first_gate))
```

# cell types test data
```{r}
cell_types_test <- c(files_to_load_names_3) %>% 
  str_remove(pattern = ".fcs") %>% 
  str_remove(pattern = "export_Specimen_..._c....._..._..._")

facs_test_data_df <- test_diff_som %>% build_fcs_data_df(cell_types = cell_types_test)

facs_test_data_df <- facs_test_data_df %>% 
  mutate(first_gate = case_when(!(cell_type %in% c("clog+", "clog-_cells-", "S1-", "S2-", "dead")) ~ "live"))
  
facs_test_data_df <- facs_test_data_df %>% 
  mutate(first_gate = if_else(is.na(first_gate), true = "not_live", false = first_gate))
```

# Split data for Cross-validation
```{r}
first_gate_table <- table(facs_data_df$first_gate)

# sort data
reseed()
facs_data_df <- facs_data_df %>% dplyr::slice(sample(nrow(facs_data_df)))

# balance data
reseed()
facs_data_subset_df <- facs_data_df %>% split(f = .$first_gate) %>% purrr::map(dplyr::slice, sample(120000)) %>% bind_rows()

# sort data
reseed()
facs_data_subset_df <- facs_data_subset_df %>% dplyr::slice(sample(nrow(facs_data_subset_df)))

k_fold <- 5

# Create 5 equally size folds
reseed()
folds <- cut(x = seq(1, nrow(facs_data_subset_df)), breaks = k_fold, labels = FALSE)
# Perform 10 fold cross validation
# Segement your data by fold using the which() function
test_indices <- purrr::map(seq_len(k_fold), function(i) which(folds == i, arr.ind = TRUE))

# Use the test and train data partitions however you desire...
test_data <- purrr::map(seq_len(k_fold), function(i) facs_data_subset_df[test_indices[[i]], ]) %>% purrr::set_names(paste0("test_", 1:k_fold))
train_data <- purrr::map(seq_len(k_fold), function(i) facs_data_subset_df[-test_indices[[i]], ]) %>% purrr::set_names(paste0("train_", 1:k_fold))
  
train_data_list <- purrr::map(seq_len(k_fold), function(i) list(measurements = train_data[[i]] %>% dplyr::select(-cell_type) %>% as.matrix(), classes_col = train_data[[i]] %>% dplyr::select(cell_type) %>% unlist()))
  
test_data_list <- purrr::map(seq_len(k_fold), function(i) list(measurements = test_data[[i]] %>% dplyr::select(-c(cell_type, first_gate)) %>% as.matrix(), classes_col = test_data[[i]] %>% dplyr::select(first_gate) %>% unlist()))
```


```{r}
diff_som_1$fs$data <- facs_data_subset_df %>% as.matrix()

k_fold <- 5

# Create 5 equally size folds
reseed()
folds <- cut(x = seq(1, nrow(diff_som_1$fs$data)), breaks = k_fold, labels = FALSE)
# Perform 10 fold cross validation
# Segement your data by fold using the which() function
test_indices <- purrr::map(seq_len(k_fold), function(i) which(folds == i, arr.ind = TRUE))

# Use the test and train data partitions however you desire...
test_data <- purrr::map(seq_len(k_fold), function(i) diff_som_1$fs$data[test_indices[[i]], ])
train_data <- purrr::map(seq_len(k_fold), function(i) diff_som_1$fs$data[-test_indices[[i]], ])

diff_som_1_test <- list(test_1 = diff_som_1,
                        test_2 = diff_som_1,
                        test_3 = diff_som_1,
                        test_4 = diff_som_1,
                        test_5 = diff_som_1)

diff_som_1_test[[1]]$fs$data <- diff_som_1_test[[1]]$fs$data %>% as.data.frame() %>% dplyr::slice(test_indices[[1]]) %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)
diff_som_1_test[[2]]$fs$data <- diff_som_1_test[[2]]$fs$data %>% as.data.frame() %>% dplyr::slice(test_indices[[2]]) %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)
diff_som_1_test[[3]]$fs$data <- diff_som_1_test[[3]]$fs$data %>% as.data.frame() %>% dplyr::slice(test_indices[[3]]) %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)
diff_som_1_test[[4]]$fs$data <- diff_som_1_test[[4]]$fs$data %>% as.data.frame() %>% dplyr::slice(test_indices[[4]]) %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)
diff_som_1_test[[5]]$fs$data <- diff_som_1_test[[5]]$fs$data %>% as.data.frame() %>% dplyr::slice(test_indices[[5]]) %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)

# purrr::map(seq_len(k_fold), function(i) diff_som_1_test[[i]]$fs$data = diff_som_1_test[[i]]$fs$data %>% as.data.frame() %>% dplyr::slice(test_indices[[i]]) %>% as.matrix() )


diff_som_1_train <- list(train_1 = diff_som_1,
                        train_2 = diff_som_1,
                        train_3 = diff_som_1,
                        train_4 = diff_som_1,
                        train_5 = diff_som_1)

diff_som_1_train[[1]]$fs$data <- diff_som_1_train[[1]]$fs$data %>% as.data.frame() %>% dplyr::slice(-test_indices[[1]]) %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)
diff_som_1_train[[2]]$fs$data <- diff_som_1_train[[2]]$fs$data %>% as.data.frame() %>% dplyr::slice(-test_indices[[2]]) %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)
diff_som_1_train[[3]]$fs$data <- diff_som_1_train[[3]]$fs$data %>% as.data.frame() %>% dplyr::slice(-test_indices[[3]]) %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)
diff_som_1_train[[4]]$fs$data <- diff_som_1_train[[4]]$fs$data %>% as.data.frame() %>% dplyr::slice(-test_indices[[4]]) %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)
diff_som_1_train[[5]]$fs$data <- diff_som_1_train[[5]]$fs$data %>% as.data.frame() %>% dplyr::slice(-test_indices[[5]]) %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)

# purrr::map(seq_len(k_fold), function(i) diff_som_1_train[[i]]$fs$data = diff_som_1_train[[i]]$fs$data[-test_indices[[i]], ])

# train_data_list <- purrr::map(seq_len(k_fold), function(i) list(measurements = train_data[[i]] %>% dplyr::select(-cell_type) %>% as.matrix(), classes_col = train_data[[i]] %>% dplyr::select(cell_type) %>% unlist()))

# test_data_list <- purrr::map(seq_len(k_fold), function(i) list(measurements = test_data[[i]] %>% dplyr::select(-cell_type) %>% as.matrix(), classes_col = test_data[[i]] %>% dplyr::select(cell_type) %>% unlist()))
```

```{r}
# process_C5 <- function(input = facs_data_df, n_clusters, ...){
process_C5 <- function(df, features, ...){  
  library(tidyverse, warn.conflicts = FALSE)
  library(caret, warn.conflicts = FALSE)
  library(C50, warn.conflicts = FALSE)
  # library(doParallel, warn.conflicts = FALSE)
  # EmbedSOM_results_df_all = process_step_1_output$EmbedSOM_results_df_all
  # EmbedSOM_results_df = process_step_1_output$EmbedSOM_results_df
  # facs_data_df = process_step_1_output$facs_data_df
  # 
  # features_all_from_SOM <- EmbedSOM_results_df_all %>% dplyr::select(dplyr::contains(".y"))
  # features_selected_from_SOM <- df %>% dplyr::select(dplyr::contains(".y"))
  # 
  
  df <- df %>% as.data.frame()
  
  features_df <- df %>% 
    dplyr::select(-c(first_gate, cell_type)) %>% 
    dplyr::select(features) %>% 
    apply(FUN = as.numeric, MARGIN = c(2)) %>% 
    as.data.frame()
  
  # features_selected_no_SOM <- df %>% dplyr::select(-c(first_gate, cell_type)) %>% dplyr::select(c(FSC.A, SSC.A, FSC.H, SSC.W, Indo.1..Violet..A))
  
  classes <- df$first_gate %>% factor()
  
  # rules_model_selected
  # rules_model <- C5.0(x = features_df, y = classes, trials = 20, rules = TRUE)
  # tree_model <- C5.0(x = features_df, y = classes, trials = 20, rules = FALSE)
  
  # rules_model_all
  rules_model <- C5.0(x = features_df, y = classes, trials = 20, rules = TRUE)
  # tree_model <- C5.0(x = features_df, y = classes, trials = 20, rules = FALSE)
  
  # rules_model_from_SOM
  # rules_model <- C5.0(x = features_selected_from_SOM, y = classes, trials = 1, rules = TRUE)
  # tree_model <- C5.0(x = features_selected_from_SOM, y = classes, trials = 1, rules = FALSE)
  
  # tree_model_from_SOM_all_features
  # tree_model <- C5.0(x = features_all_from_SOM, y = classes, trials = 1, rules = FALSE)
  # rules_model <- C5.0(x = features_all_from_SOM, y = classes, trials = 1, rules = TRUE)

    return(rules_model)
}


df_1 <- train_data[[1]] %>% 
  as.data.frame() %>% 
  dplyr::select(-c(first_gate, cell_type)) %>% 
  dplyr::select(c(1:18)) %>% 
  apply(FUN = as.numeric, MARGIN = c(2)) %>% 
  as.data.frame()
```

# No SOM - use data directly
```{r}
registerDoParallel(cores = 5)

# 1945.741 sec elapsed
tic()
rules_models <- foreach(i = seq_along(train_data)) %dopar% process_C5(df = train_data[[i]], features = c(1:18))
toc()

# tic()
# rules_models <- train_data %>% purrr::map(function(x) x %>% process_C5(features = c(1:18)))
# toc()

# 66784.289 sec elapsed
# tic()
# rules_models <- train_data %>% purrr::map(function(x) x %>% process_C5())
# toc()
# rules_models %>% saveRDS(file = "~/Downloads/rules_models_2020_06_09.rds")
```

# save

```{r}
summary(rules_models[[3]])

rules_models_results <- rules_models %>% purrr::map(summary) %>%  purrr::map(capture.output)

purrr::map(seq_along(rules_models_results), function(i) rules_models_results[[i]] %>% cat(file = paste0("~/Downloads/rules_models_results_", i, ".txt")))
```


Validate
```{r}
# 48.722 sec elapsed
tic()
predictions_validation <- purrr::map(seq_along(rules_models_results), function(i) predict(rules_models[[i]], newdata = test_data_list[[i]]$measurements))
toc()
```

# Confusion matrix
```{r}
confusion_matrix_results <- purrr::map(seq_along(predictions_validation), function(i) caret::confusionMatrix(predictions_validation[[i]], test_data_list[[i]]$classes_col %>% factor()))

mean_accuracy <- confusion_matrix_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% mean()

sd_accuracy <- confusion_matrix_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% sd()

mean_kappa <- confusion_matrix_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Kappa"]) %>% unlist() %>% mean()
```

# Test if cross-validation really works with new data
```{r}
test_df <- facs_test_data_df %>% dplyr::select(-c(cell_type, first_gate))

test_results <- purrr::map(seq_along(rules_models_results), function(i) predict(rules_models[[i]], newdata = test_df))

confusion_matrix_test_results <- purrr::map(seq_along(test_results), function(i) caret::confusionMatrix(test_results[[i]], facs_test_data_df$first_gate %>% factor()))

confusion_matrix_test_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% mean()
confusion_matrix_test_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Kappa"]) %>% unlist() %>% mean()
```





# No SOM - Selected features
```{r}
registerDoParallel(cores = 5)

# 184.415 sec elapsed
tic()
rules_models_selected_features <- foreach(i = seq_along(train_data)) %dopar% process_C5(df = train_data[[i]], features = c("FSC.A", "SSC.A", "FSC.H", "SSC.W", "Indo.1..Violet..A"))
toc()

# 67283.483 sec elapsed
# tic()
# rules_models_selected_features <- train_data %>% purrr::map(dplyr::select, c(FSC.A, SSC.A, FSC.H, SSC.W, Indo.1..Violet..A)) %>% purrr::map(function(x) x %>% process_C5(n_clusters = 6))
# toc()

# rules_models_selected_features %>% saveRDS(file = "~/Downloads/rules_models_selected_features_results_2020_06_10.rds")
```

```{r}
rules_models_selected_features_results <- rules_models_selected_features %>% purrr::map(summary) %>%  purrr::map(capture.output)

purrr::map(seq_along(rules_models_selected_features_results), function(i) rules_models_selected_features_results[[i]] %>% cat(file = paste0("~/Downloads/rules_models_selected_features_results_", i, ".txt")))
```

# Validate
```{r}
# 12.661 sec elapsed
tic()
selected_features_predictions_validation <- purrr::map(seq_along(rules_models_selected_features_results), function(i) predict(rules_models_selected_features[[i]], newdata = test_data_list[[i]]$measurements))
toc()
```

# Confusion matrix
```{r}
confusion_matrix_selected_features_results <- purrr::map(seq_along(selected_features_predictions_validation), function(i) caret::confusionMatrix(selected_features_predictions_validation[[i]], test_data_list[[i]]$classes_col %>% factor()))

mean_selected_features_accuracy <- confusion_matrix_selected_features_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% mean()

sd_selected_features_accuracy <- confusion_matrix_selected_features_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% sd()

confusion_matrix_selected_features_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Kappa"]) %>% unlist() %>% mean()
```

# Test if cross-validation really works with new data
```{r}
test_df <- facs_test_data_df %>% dplyr::select(-c(cell_type, first_gate))

test_selected_features_results <- purrr::map(seq_along(rules_models_results), function(i) predict(rules_models_selected_features[[i]], newdata = test_df))

confusion_matrix_test_selected_features_results <- purrr::map(seq_along(test_selected_features_results), function(i) caret::confusionMatrix(test_selected_features_results[[i]], facs_test_data_df$first_gate %>% factor()))

confusion_matrix_test_selected_features_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% mean()

confusion_matrix_test_selected_features_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Kappa"]) %>% unlist() %>% mean()
```






```{r}
# diff_som <- diff_som_1_train[[1]]

perform_SOM <- function(diff_som, facs_data_df, columns_to_use = c("FSC.A", "FSC.H", "SSC.A", "SSC.W", "Indo.1..Violet..A"), importance = c(1, 1, 1, 1, 1), normalize_weights = TRUE) {

  diff_som = diff_som
  
  # calculate optimal grid
  vesanto_dim <- diff_som$fs$data %>%
    nrow() %>%
    sqrt() %>%
    "*"(5)
  
  grid_dim <- vesanto_dim %>%
    sqrt() %>%
    ceiling() %>%
    "+"(20)

  # SOM
  reseed()
  ds_step_1 <- diff_som %>% Embed(xdim = grid_dim, ydim = grid_dim, colsToUse = columns_to_use, importance = importance, nclust = grid_dim * grid_dim, rlen = 5)

  # Optional to normalize weights
  if (normalize_weights) {
    ds_step_1$map$codes <- purrr::map(seq_len(ncol(ds_step_1$map$codes)), function(i) {ds_step_1$map$codes[, i] %>% bestNormalize::orderNorm() %>% .[["x.t"]] %>% scale_to_range_0_1()}) %>% purrr::set_names(colnames(ds_step_1$map$codes)) %>% bind_cols()
  }

  grid_clustering_df <- ds_step_1$map$grid %>%
    as.data.frame() %>%
    rownames_to_column() %>%
    mutate(cluster = ds_step_1$clust) %>%
    mutate(rowname = rowname %>% as.numeric()) %>%
    bind_cols(as.data.frame(ds_step_1$map$codes))

  EmbedSOM_results_df <- ds_step_1$fs$data %>%
    as.data.frame() %>%
    mutate(mapping = ds_step_1$map$mapping[, 1]) %>%
    left_join(grid_clustering_df, by = c("mapping" = "rowname")) %>%
    mutate(first_gate = facs_data_df$first_gate)

  output <- list(EmbedSOM_results_df = EmbedSOM_results_df,
                 diff_som = ds_step_1)
  
  return(output)
}

```


# SOM - Selected features
```{r}
registerDoParallel(cores = 5)
# 175.813 sec elapsed - serial
tic()
# EmbedSOM_results_train <- purrr::map(seq_along(diff_som_1_train), function(i) perform_SOM(diff_som = diff_som_1_train[[i]], facs_data_subset_df[-test_indices[[i]], ], columns_to_use = c("FSC.A", "FSC.H", "SSC.A", "SSC.W", "Indo.1..Violet..A"), importance = c(1, 1, 1, 1, 1), normalize_weights = TRUE))
EmbedSOM_results_train <- purrr::map(seq_along(diff_som_1_train), function(i) perform_SOM(diff_som = diff_som_1_train[[i]], facs_data_subset_df[-test_indices[[i]], ], columns_to_use = c("FSC.A", "FSC.H", "SSC.A", "SSC.W", "Indo.1..Violet..A"), importance = c(1, 1, 1, 1, 1), normalize_weights = FALSE))
toc()

# 40.863 sec elapsed
tic()
# EmbedSOM_results_test <- purrr::map(seq_along(diff_som_1_test), function(i) perform_SOM(diff_som = diff_som_1_test[[i]], facs_data_subset_df[test_indices[[i]], ], columns_to_use = c("FSC.A", "FSC.H", "SSC.A", "SSC.W", "Indo.1..Violet..A"), importance = c(1, 1, 1, 1, 1), normalize_weights = TRUE))
EmbedSOM_results_test <- purrr::map(seq_along(diff_som_1_test), function(i) perform_SOM(diff_som = diff_som_1_test[[i]], facs_data_subset_df[test_indices[[i]], ], columns_to_use = c("FSC.A", "FSC.H", "SSC.A", "SSC.W", "Indo.1..Violet..A"), importance = c(1, 1, 1, 1, 1), normalize_weights = FALSE))
toc()
```

# C5.0 on SOM
```{r}
process_C5_from_SOM <- function(df, features, ...){  
  library(tidyverse, warn.conflicts = FALSE)
  library(caret, warn.conflicts = FALSE)
  library(C50, warn.conflicts = FALSE)
  # EmbedSOM_results_df_all = process_step_1_output$EmbedSOM_results_df_all
  # EmbedSOM_results_df = process_step_1_output$EmbedSOM_results_df
  # facs_data_df = process_step_1_output$facs_data_df
  # 
  # features_all_from_SOM <- EmbedSOM_results_df_all %>% dplyr::select(dplyr::contains(".y"))
  weights_df <- df %>% dplyr::select(features)
  # 
  # features_all_no_SOM <- facs_data_df %>% dplyr::select(-c(first_gate, cell_type))
  # features_selected_no_SOM <- facs_data_df %>% dplyr::select(-c(first_gate, cell_type)) %>% dplyr::select(c(FSC.A, SSC.A, FSC.H, SSC.W, Indo.1..Violet..A))
  
  classes <- df$first_gate %>% factor()

  # rules_model_selected
  # rules_model <- C5.0(x = features_df, y = classes, trials = 20, rules = TRUE)
  # tree_model <- C5.0(x = features_df, y = classes, trials = 20, rules = FALSE)
  
  # rules_model_all
  # rules_model <- C5.0(x = features_df, y = classes, trials = 20, rules = TRUE)
  # tree_model <- C5.0(x = features_df, y = classes, trials = 20, rules = FALSE)
  
  # rules_model_from_SOM
  rules_model <- C5.0(x = weights_df, y = classes, trials = 1, rules = TRUE)
  # tree_model <- C5.0(x = weights_df, y = classes, trials = 1, rules = FALSE)
  
  # tree_model_from_SOM_all_features
  # tree_model <- C5.0(x = weights_df, y = classes, trials = 1, rules = FALSE)
  # rules_model <- C5.0(x = weights_df, y = classes, trials = 1, rules = TRUE)
  return(rules_model)
}

# 78.039 sec elapsed - serial
tic()
# rules_models_SOM_selected_features <- EmbedSOM_results_train %>% purrr::map(function(x) x %>% .[["EmbedSOM_results_df"]] %>% process_C5(dplyr::contains(".y")))
# rules_models_SOM_selected_features <- EmbedSOM_results_train %>% purrr::map(function(x) x %>% .[["diff_som"]] %>% .[["map"]] %>% .[["codes"]] %>% process_C5(n_clusters = 6))
# rules_models_SOM_selected_features <- purrr::map(seq_along(EmbedSOM_results_train), function(i) EmbedSOM_results_train[[i]]$EmbedSOM_results_df %>% process_C5(n_clusters = 6))

rules_models_SOM_selected_features <- foreach(i = seq_along(EmbedSOM_results_train)) %dopar% process_C5_from_SOM(df = EmbedSOM_results_train[[i]]$EmbedSOM_results_df, features = dplyr::contains(".y"))
toc()

# EmbedSOM_results_train[[1]]$diff_som$map$codes
# EmbedSOM_results_train[[1]]$EmbedSOM_results_df

rules_models_SOM_selected_features %>% saveRDS(file = "~/Downloads/rules_models_SOM_selected_features.rds")
```


# Validate
```{r}
# 2.862 sec elapsed
tic()
selected_features_SOM_predictions_validation <- purrr::map(seq_along(rules_models_SOM_selected_features), function(i) predict(rules_models_SOM_selected_features[[i]], newdata = EmbedSOM_results_test[[i]]$EmbedSOM_results_df %>% dplyr::select(dplyr::contains(".y"))))
toc()
```

# Confusion matrix
```{r}
confusion_matrix_selected_features_SOM_results <- purrr::map(seq_along(selected_features_SOM_predictions_validation), function(i) caret::confusionMatrix(selected_features_SOM_predictions_validation[[i]], EmbedSOM_results_test[[i]]$EmbedSOM_results_df$first_gate %>% factor()))

mean_selected_features_SOM_accuracy <- confusion_matrix_selected_features_SOM_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% mean()

sd_selected_features_SOM_accuracy <- confusion_matrix_selected_features_SOM_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% sd()

mean_selected_features_SOM_kappa <- confusion_matrix_selected_features_SOM_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Kappa"]) %>% unlist() %>% mean()
```

# Test if cross-validation really works with new data
```{r}
# facs_test_data_df
# test_diff_som

# test_diff_som$fs$data <- test_diff_som$fs$data %>% as.data.frame() %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)

test_diff_som$fs$data <- test_diff_som$fs$data %>% 
  as.data.frame() %>% 
  purrr::set_names(str_remove(string = colnames(test_diff_som$fs$data), pattern = "FJComp.")) %>% 
  as.matrix()

test_df <- facs_test_data_df %>% dplyr::select(-c(cell_type, first_gate))

EmbedSOM_results_testing <- perform_SOM(diff_som = test_diff_som, facs_data_df = facs_test_data_df, columns_to_use = c("FSC.A", "FSC.H", "SSC.A", "SSC.W", "Indo.1..Violet..A"), importance = c(1, 1, 1, 1, 1), normalize_weights = TRUE)

test_selected_features_SOM_results <- purrr::map(seq_along(rules_models_SOM_selected_features), function(i) predict(rules_models_SOM_selected_features[[i]], newdata = EmbedSOM_results_testing$EmbedSOM_results_df %>% dplyr::select(dplyr::contains(".y"))))

confusion_matrix_test_selected_SOM_features_results <- purrr::map(seq_along(test_selected_features_SOM_results), function(i) caret::confusionMatrix(test_selected_features_SOM_results[[i]], facs_test_data_df$first_gate %>% factor()))

confusion_matrix_test_selected_SOM_features_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% mean()

confusion_matrix_test_selected_SOM_features_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Kappa"]) %>% unlist() %>% mean()
```








# SOM - all features
```{r}
# 344.531 sec elapsed
# 360.017 sec elapsed
tic()
# EmbedSOM_all_features_results_train <- purrr::map(seq_along(diff_som_1_train), function(i) perform_SOM(diff_som = diff_som_1_train[[i]], facs_data_subset_df[-test_indices[[i]], ], columns_to_use = c(1:18), importance = rep(1, 18), normalize_weights = TRUE))
EmbedSOM_all_features_results_train <- purrr::map(seq_along(diff_som_1_train), function(i) perform_SOM(diff_som = diff_som_1_train[[i]], facs_data_subset_df[-test_indices[[i]], ], columns_to_use = c(1:18), importance = rep(1, 18), normalize_weights = FALSE))
toc()

# 65.526 sec elapsed
# 67.667 sec elapsed
tic()
# EmbedSOM_all_features_results_test <- purrr::map(seq_along(diff_som_1_test), function(i) perform_SOM(diff_som = diff_som_1_test[[i]], facs_data_subset_df[test_indices[[i]], ], columns_to_use = c(1:18), importance = rep(1, 18), normalize_weights = TRUE))
EmbedSOM_all_features_results_test <- purrr::map(seq_along(diff_som_1_test), function(i) perform_SOM(diff_som = diff_som_1_test[[i]], facs_data_subset_df[test_indices[[i]], ], columns_to_use = c(1:18), importance = rep(1, 18), normalize_weights = FALSE))
toc()
```

# C5.0 on SOM
```{r}
registerDoParallel(cores = 5)
# 104.557 sec elapsed
tic()
rules_models_SOM_all_features <- EmbedSOM_all_features_results_train %>% purrr::map(function(x) x %>% .[["EmbedSOM_results_df"]] %>% process_C5_from_SOM(features = dplyr::contains(".y")))

# rules_models_SOM_all_features <- foreach(i = seq_along(EmbedSOM_all_features_results_train)) %dopar% process_C5_from_SOM(df = EmbedSOM_all_features_results_train[[i]]$EmbedSOM_results_df, features = dplyr::contains(".y"))
toc()

# EmbedSOM_results_train[[1]]$diff_som$map$codes
# EmbedSOM_results_train[[1]]$EmbedSOM_results_df

# rules_models_SOM_all_features %>% saveRDS(file = "~/Downloads/rules_models_SOM_all_features.rds")
```


# Validate
```{r}
# 8.65 sec elapsed
tic()
all_features_SOM_predictions_validation <- purrr::map(seq_along(rules_models_SOM_all_features), function(i) predict(rules_models_SOM_all_features[[i]], newdata = EmbedSOM_all_features_results_test[[i]]$EmbedSOM_results_df %>% dplyr::select(dplyr::contains(".y"))))
toc()
```

# Confusion matrix
```{r}
confusion_matrix_all_features_SOM_results <- purrr::map(seq_along(all_features_SOM_predictions_validation), function(i) caret::confusionMatrix(all_features_SOM_predictions_validation[[i]], EmbedSOM_all_features_results_test[[i]]$EmbedSOM_results_df$first_gate %>% factor()))

mean_all_features_SOM_accuracy <- confusion_matrix_all_features_SOM_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% mean()

sd_all_features_SOM_accuracy <- confusion_matrix_all_features_SOM_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% sd()

mean_all_features_SOM_kappa <- confusion_matrix_all_features_SOM_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Kappa"]) %>% unlist() %>% mean()
```

# Test if cross-validation really works with new data
```{r}
# facs_test_data_df
# test_diff_som

# test_diff_som$fs$data <- test_diff_som$fs$data %>% as.data.frame() %>% dplyr::select(-c("cell_type", "first_gate")) %>% as.matrix() %>% apply(MARGIN = c(2), as.numeric)

test_diff_som$fs$data <- test_diff_som$fs$data %>% 
  as.data.frame() %>% 
  purrr::set_names(str_remove(string = colnames(test_diff_som$fs$data), pattern = "FJComp.")) %>% 
  as.matrix()

test_df <- facs_test_data_df %>% dplyr::select(-c(cell_type, first_gate))

EmbedSOM_all_features_results_testing <- perform_SOM(diff_som = test_diff_som, facs_data_df = facs_test_data_df, columns_to_use = c(1:18), importance = rep(1, 18), normalize_weights = TRUE)

test_all_features_SOM_results <- purrr::map(seq_along(rules_models_SOM_all_features), function(i) predict(rules_models_SOM_all_features[[i]], newdata = EmbedSOM_all_features_results_testing$EmbedSOM_results_df %>% dplyr::select(dplyr::contains(".y"))))

confusion_matrix_test_all_SOM_features_results <- purrr::map(seq_along(test_all_features_SOM_results), function(i) caret::confusionMatrix(test_all_features_SOM_results[[i]], facs_test_data_df$first_gate %>% factor()))






mean_selected_all_features_SOM_accuracy <- confusion_matrix_test_all_SOM_features_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% mean()

sd_selected_all_features_SOM_accuracy <- confusion_matrix_test_all_SOM_features_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Accuracy"]) %>% unlist() %>% sd()

mean_selected_all_features_SOM_kappa <- confusion_matrix_test_all_SOM_features_results %>% purrr::map(function(x) x %>% .[["overall"]] %>% .["Kappa"]) %>% unlist() %>% mean()
```

```{r}

```


